---
title: "How Credibility of Information About Gender Affirming Care Impacts Support for Legislation"
subtitle: "DACSS 602 (Fall 2024)"
author: Maura Anish, Kendrick Backmon, Avery Bacon, and Jules Tucher 
format: 
  html:
    toc: true
    toc-depth: 2
    toc-title: Contents
    toc-location: left
    html-math-method: katex
    theme: flatly
    smooth-scroll: true
    link-external-icon: true
    link-external-newwindow: true
    citations-hover: true
    footnotes-hover: true
    font-size: 80%
---

<!--
ATTENTION: Please submit both your .qmd file and the corresponding ".html" file after rendering.
-->


<!--
Replace the placeholders (title and author) in the yaml above with your own.
-->


# Introduction
<!--
What is your experiment about (4-5 sentences).
-->

# Methodology

In order to answer our research question, we designed a 2x2 factorial between-subjects experiment with 4 treatment groups. We tested two nominal independent variables: credibility of information and inclusion of a statistic supporting the information. Credibility of information is provided by including a citation by well-known and credible source, in this case the Mayo Clinic. The statistic included to support the claim that puberty blockers can reduce risk of suicidality was "in some cases by over 70%". The four treatment graphics, including the full text of the infographic, are provided in Figure 1.

The experiment's outcome variable is the degree of support for legislation banning gender-affirming care for minors. This was measured using a 5-point Likert type scale: Strongly Oppose (1), Slightly Oppose (2), Neutral (3), Slightly Approve (4), and Strongly Approve (5). 

To complete the experimental design, respondents first completed an informed consent form and were asked several demographic questions. Third variables of particular interest to our research are age, gender, region, and political affiliation. Furthermore, survey respondents were asked a prior knowledge assessment to determine prior bias and understanding of the topic: "True or False: Puberty blockers are a safe and reversible treatment when prescribed by medical professionals to treat gender dysphoria in transgender minors."

A total of 220 survey responses were collected. 
<!--
Describe your experiment design, including: (1) experiment design, (2) treatments, (3) outcome variable(s), (4) sample size, and (5) data cleaning procedures (if applicable) [8-10 sentences]
-->

# Analysis

We could use either factorial ANOVA or Ordered Logistic Regression to explore the results of our survey. Because the outcome variable is a Likert-type ordinal variable and not continuous, it does not meet the continuity assumption for factorial ANOVA. Also, Ordered Logistic Regression would allow for us to control for age, gender, region, and political affiliation.

<!--
Which statistical test is used, and why? [3-4 sentences]
-->

<!--
Run the chosen test! To do so: (1) import the libraries you need, (2) import your data, (3) double-check the names of your IV(s) and DV, and (4) run the test. For information on the libraries and R syntax, refer to see Lecture 10.
-->

```{r}
# import libraries
library(tidyverse)
library(ggplot2)


# import data
g1_data <- read.csv("./fa24_g1.csv") %>%
  mutate(IV_org = as.factor(ifelse(IV_org == "no", "Non-credible", "Credible")),
         IV_info = as.factor(ifelse(IV_info == "no", "No statistic", "Statistic")),
         prior_knowledge = factor(g1q1, levels=c("True", "False", "I don't know")),
         outcome = factor(g1q2,
                          levels=c("Strongly oppose", "Slightly oppose", "Neutral", "Slightly support", "Strongly support"),
                          ordered=TRUE),
         manipulation_check = factor(g1q3,
                                     levels=c("Do not support", "Support", "Undecided")))
summary(g1_data)


# recreate expected results plot
summary_table <- g1_data %>%
  mutate(treatment = paste0(IV_org, ",\n", IV_info)) %>%
  select(treatment, outcome, prior_knowledge) %>%
  group_by(treatment, outcome, prior_knowledge) %>%
  summarize(value = n())
ggplot(summary_table, aes(fill=outcome, y=value, x=treatment)) + 
    geom_bar(position="fill", stat="identity")  +
    facet_wrap(~ prior_knowledge) + 
    labs(x = "Exposure", y = "Proportion of Sample") + 
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45))


    scale_fill_manual(values = c("#881c1c", "#ffc72c", "#212721"),
                     name = "Orientation to\nLegislation",
                     labels = c("Against", "Neutral", "Support"))
    

# data cleaning steps

# explore consistency with manipulation check, throw out any inconsistent results

# two-way ANOVA

# ordered logistic regression

```




# Results
<!--
Report and interpret the test results. For information on what should be included, refer to lecture 10. Also add at least one plot.
-->




# Findings
<!--
Findings: Report the statistic you obtained along with the information to decide whether the null hypothesis can be rejected or not. Use a plot and/or table with a proper label.
-->



# Discussion
<!--
What are the implications of the study? What are possible explanations if the results do not align with your hypothesis? Any suggestions for future studies?
-->



### For your information
<!--
Some example code for different plots. Below is only for your information. Submit your report after removing this section.
-->

```{r}
# the built-in R data set named ToothGrowth:
# data from a study evaluating the effect of Vitamin C on tooth growth in Guinea pigs.
# The experiment has been performed on 60 pigs, where each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC). Tooth length was measured and a sample of the data is shown below
my_data = ToothGrowth
str(my_data)

# Convert dose as a factor and recode the levels
# as D0.5, D1, D2
my_data$dose <- factor(my_data$dose,
                       levels = c(0.5, 1, 2),
                       labels = c("D0.5", "D1", "D2"))
head(my_data)

# frequency table:
table(my_data$supp, my_data$dose)
```


```{r}
# Line plots with multiple groups
# plot tooth length ("len") by groups ("dose")
# color box plot by a second group: "supp"
# add error bars: mean_se
library(ggpubr)
ggline(my_data, x = "dose", y = "len", color = "supp",
       add = c("mean_se", "dotplot"),
       palette = c("#00AFBB", "#E7B800"))

# OR
# using R base graphs:
boxplot(len ~ supp * dose, data = my_data, frame = FALSE,
        col = c("#00AFBB", "#E7B800"), ylab = "Tooth Length")

# Another example: Two-way interaction plot
interaction.plot(x.factor = my_data$dose, trace.factor = my_data$supp, 
                 # x.factor: the factor to be plotted on x axis
                 # trace.factor: the factor to be plotted as lines
                 # response: a numeric variable giving the response
                 response = my_data$len, fun = mean,
                 # type: the type of plot (p: point, l: line, b: both point and line)
                 type = "b", legend = TRUE,
                 xlab = "Dose", ylab = "Tooth Length",
                 pch = c(1, 19), col = c("#00AFBB", "#E7B800"))
```

